Untitled

### 2023-02-14 Manuscript/Lanuza_et_al_2022.Rmd
Edits on code to run without the last col of nectar conc that is not finally added

2. Here, using a global dataset of 16 reproductive traits for 1,506 species, we investigate the reproductive spectrum of flowering plants to identify how it shapes interactions with pollinators.

@@

We selected a total of 19 quantitative and categorical functional traits that included reproductive traits and traits related with plant form and size (see table 1). Reproductive traits consisted of those related to floral and reproductive biology: (i) floral traits included those related to the size of floral organs (e.g., style length), floral display size (e.g., flower number) and floral rewards (e.g., pollen quantity), while (ii) reproductive biology traits represented the plant reproductive system (i.e., the breeding, mating and compatibility system). The remaining traits were related to plant size and form (i.e., plant height, lifespan and life form), which are commonly used to characterise the fast-slow continuum of plant trait variation [@salguero2016]. For each plant species, we undertook an extensive literature and online search for all traits across a wide range of resources (plant databases, online floras, books, journals and images). From a matrix comprising a total of 28,614 possible cells (considering all traits and plant species: 19 columns × 1,506 species), we were able to fill 23,354 cells (81.6% of the dataset, see figure S2 for missing value information for each trait). An extended description of each trait and how it was obtained can be found in Appendix S1.

@@

Trait <- c("Plant height", "Flower width", "Flower length", "Inflorescence width", "Style length", "Ovules per flower",
           "Flowers per plant", "Microliters of nectar", "Milligrams of nectar", "Nectar concentration", "Pollen grains per flower", "Autonomous selfing (fruit set)")

@@

T.range <- c("0.01-30m", "0.50-205mm", "0.20-195mm", "0.80-300mm", "0-138mm", "1-25000", "1-1.8 x $10^5$", "0-160$\\mu$l", "0-6.1mg", "0-100$\\%$", "13-2 x $10^8$", "0-100$\\%$")

@@

kable(d, booktabs = T, linesep = "\\addlinespace", escape = FALSE,caption = "\\textbf{Table 1.} Quantitative and categorical traits used in this study.") %>%
  kable_styling(full_width = F,font_size = 12) %>% row_spec(0, bold=T) %>% pack_rows("Quantitative", 1, 12) %>%
pack_rows("Categorical", 13, 19)

@@

We found a strong phylogenetic signal (*P* \< 0.01) for most quantitative traits (table S5). The traits that showed the highest phylogenetic signal were ovule number ($\lambda$ = 1), pollen grains per flower ($\lambda$ = 1) and plant height ($\lambda$ = 0.96), followed by flower length ($\lambda$ = 0.75), flower width ($\lambda$ = 0.73), number of flowers per plant ($\lambda$ = 0.69) and nectar concentration ($\lambda$ = 0.65). The traits that had a moderate phylogenetic signal included inflorescence width ($\lambda$ = 0.57), style length ($\lambda$ = 0.49) and autonomous selfing ($\lambda$ = 0.34). Finally, microliters of nectar per flower showed the lowest phylogenetic signal of all the traits we measured ($\lambda$ = 0.14).

>>>>

2. Here, using a global dataset of 18 reproductive traits for 1,506 species, we investigate the reproductive spectrum of flowering plants to identify how it shapes interactions with pollinators.

@@

We selected a total of 18 quantitative and categorical functional traits that included reproductive traits and traits related with plant form and size (see table 1). Reproductive traits consisted of those related to floral and reproductive biology: (i) floral traits included those related to the size of floral organs (e.g., style length), floral display size (e.g., flower number) and floral rewards (e.g., pollen quantity), while (ii) reproductive biology traits represented the plant reproductive system (i.e., the breeding, mating and compatibility system). The remaining traits were related to plant size and form (i.e., plant height, lifespan and life form), which are commonly used to characterise the fast-slow continuum of plant trait variation [@salguero2016]. For each plant species, we undertook an extensive literature and online search for all traits across a wide range of resources (plant databases, online floras, books, journals and images). From a matrix comprising a total of 27,108 possible cells (considering all traits and plant species: 18 columns × 1,506 species), we were able to fill 22,796 cells (84.1% of the dataset, see figure S2 for missing value information for each trait). An extended description of each trait and how it was obtained can be found in Appendix S1.

@@

Trait <- c("Plant height", "Flower width", "Flower length", "Inflorescence width", "Style length", "Ovules per flower",
           "Flowers per plant", "Microliters of nectar", "Milligrams of nectar", "Pollen grains per flower", "Autonomous selfing (fruit set)")

@@

T.range <- c("0.01-30m", "0.50-205mm", "0.20-195mm", "0.80-300mm", "0-138mm", "1-25000", "1-1.8 x $10^5$", "0-160$\\mu$l", "0-6.1mg", "13-2 x $10^8$", "0-100$\\%$")

@@

kable(d, booktabs = T, linesep = "\\addlinespace", escape = FALSE,caption = "\\textbf{Table 1.} Quantitative and categorical traits used in this study.") %>%
  kable_styling(full_width = F,font_size = 12) %>% row_spec(0, bold=T) %>% pack_rows("Quantitative", 1, 11) %>%
pack_rows("Categorical", 12, 18)

@@

We found a strong phylogenetic signal (*P* \< 0.01) for most quantitative traits (table S5). The traits that showed the highest phylogenetic signal were ovule number ($\lambda$ = 1), pollen grains per flower ($\lambda$ = 1) and plant height ($\lambda$ = 0.96), followed by flower length ($\lambda$ = 0.75), flower width ($\lambda$ = 0.73) and number of flowers per plant ($\lambda$ = 0.69). The traits that had a moderate phylogenetic signal included inflorescence width ($\lambda$ = 0.57), style length ($\lambda$ = 0.49) and autonomous selfing ($\lambda$ = 0.34). Finally, microliters of nectar per flower showed the lowest phylogenetic signal of all the traits we measured ($\lambda$ = 0.14).

### 2023-02-14 Manuscript/Lanuza_et_al_2022.tex
Edits on code to run without the last col of nectar conc that is not finally added

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Globally, plants display enormous variation in life history strategies and trait combinations. However, evidence suggests that evolutionary and physiological constraints limit the number of plant ecological strategies. Although there have been recent advances in understanding correlations among plant traits, reproductive traits are rarely considered, despite their key role in shaping plant life-history strategies and interactions with pollinators.
\item
  Here, using a global dataset of 16 reproductive traits for 1,506 species, we investigate the reproductive spectrum of flowering plants to identify how it shapes interactions with pollinators.
\item
  We show that over 50\% of all trait variation is explained by the first two reproductive axes, which represent the negative correlation between flower number and flower size, and the negative correlation between autonomous selfing and floral display size. In addition, these reproductive axes were associated with different numbers of interacting partners and visitation rates of the distinct pollinator guilds. However, reproductive axes explain a relatively small amount of variance in pollinator interactions highlighting the need to incorporate other factors along with reproductive traits to fully explain large scale patterns of plant-pollinator interactions.
\item
  Our study identifies the major reproductive trait correlations in flowering plants and their role in shaping networks of plant-pollinator interactions at a global scale.
\end{enumerate}

@@

We selected a total of 19 quantitative and categorical functional traits that included reproductive traits and traits related with plant form and size (see table 1). Reproductive traits consisted of those related to floral and reproductive biology: (i) floral traits included those related to the size of floral organs (e.g., style length), floral display size (e.g., flower number) and floral rewards (e.g., pollen quantity), while (ii) reproductive biology traits represented the plant reproductive system (i.e., the breeding, mating and compatibility system). The remaining traits were related to plant size and form (i.e., plant height, lifespan and life form), which are commonly used to characterise the fast-slow continuum of plant trait variation (\protect\hyperlink{ref-salguero2016}{Salguero-Gómez et al., 2016}). For each plant species, we undertook an extensive literature and online search for all traits across a wide range of resources (plant databases, online floras, books, journals and images). From a matrix comprising a total of 28,614 possible cells (considering all traits and plant species: 19 columns × 1,506 species), we were able to fill 23,354 cells (81.6\% of the dataset, see figure S2 for missing value information for each trait). An extended description of each trait and how it was obtained can be found in Appendix S1.

@@

\caption{\label{tab:unnamed-chunk-1}\textbf{Table 1.} Quantitative and categorical traits used in this study.}
\centering
\fontsize{12}{14}\selectfont
\begin{tabular}[t]{ll}
\toprule
\textbf{Traits} & \textbf{Numerical range/Categories}\\
\midrule
\addlinespace[0.3em]
\multicolumn{2}{l}{\textbf{Quantitative}}\\
\hspace{1em}Plant height & 0.01-30m\\
\hspace{1em}Flower width & 0.50-205mm\\
\hspace{1em}Flower length & 0.20-195mm\\
\hspace{1em}Inflorescence width & 0.80-300mm\\
\hspace{1em}Style length & 0-138mm\\
\hspace{1em}Ovules per flower & 1-25000\\
\hspace{1em}Flowers per plant & 1-1.8 x $10^5$\\
\hspace{1em}Microliters of nectar & 0-160$\mu$l\\
\hspace{1em}Milligrams of nectar & 0-6.1mg\\
\hspace{1em}Nectar concentration & 0-100$\%$\\
\hspace{1em}Pollen grains per flower & 13-2 x $10^8$\\
\hspace{1em}Autonomous selfing (fruit set) & 0-100$\%$\\
\addlinespace[0.3em]
\multicolumn{2}{l}{\textbf{Categorical}}\\
\hspace{1em}Lifepan & Short-lived and perennial\\
\hspace{1em}Life form & Herb, shrub and tree\\
\hspace{1em}Flower shape & \makecell[l]{Brush, campanulate, capitulum,\\ open, papilionaceous and tube}\\
\hspace{1em}Flower symmetry & Actinomorphic and zygomorphic\\
\hspace{1em}Nectar & Presence and absence\\
\hspace{1em}Compatibility system & \makecell[l]{Self-incompatible, partially self-compabtile \\ and self-compatible}\\
\hspace{1em}Breeding system & Hermaphrodite, monoecious and dioecious\\
\bottomrule
\end{tabular}
\end{table}

>>>>

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Globally, plants display enormous variation in life history strategies and trait combinations. However, evidence suggests that evolutionary and physiological constraints limit the number of plant ecological strategies. Although there have been recent advances in understanding correlations among plant traits, reproductive traits are rarely considered, despite their key role in shaping plant life-history strategies and interactions with pollinators.
\item
  Here, using a global dataset of 18 reproductive traits for 1,506 species, we investigate the reproductive spectrum of flowering plants to identify how it shapes interactions with pollinators.
\item
  We show that over 50\% of all trait variation is explained by the first two reproductive axes, which represent the negative correlation between flower number and flower size, and the negative correlation between autonomous selfing and floral display size. In addition, these reproductive axes were associated with different numbers of interacting partners and visitation rates of the distinct pollinator guilds. However, reproductive axes explain a relatively small amount of variance in pollinator interactions highlighting the need to incorporate other factors along with reproductive traits to fully explain large scale patterns of plant-pollinator interactions.
\item
  Our study identifies the major reproductive trait correlations in flowering plants and their role in shaping plant-pollinator interactions at a macro-ecological scale. These findings emphasize the importance of considering reproductive traits in the global spectrum of plant form and function, and the need to explore beyond floral morphological traits to broaden our understanding of plant-pollinator interactions.
\end{enumerate}

@@

We selected a total of 18 quantitative and categorical functional traits that included reproductive traits and traits related with plant form and size (see table 1). Reproductive traits consisted of those related to floral and reproductive biology: (i) floral traits included those related to the size of floral organs (e.g., style length), floral display size (e.g., flower number) and floral rewards (e.g., pollen quantity), while (ii) reproductive biology traits represented the plant reproductive system (i.e., the breeding, mating and compatibility system). The remaining traits were related to plant size and form (i.e., plant height, lifespan and life form), which are commonly used to characterise the fast-slow continuum of plant trait variation (\protect\hyperlink{ref-salguero2016}{Salguero-Gómez et al., 2016}). For each plant species, we undertook an extensive literature and online search for all traits across a wide range of resources (plant databases, online floras, books, journals and images). From a matrix comprising a total of 28,614 possible cells (considering all traits and plant species: 19 columns × 1,506 species), we were able to fill 23,354 cells (81.6\% of the dataset, see figure S2 for missing value information for each trait). An extended description of each trait and how it was obtained can be found in Appendix S1.

@@

\caption{\label{tab:unnamed-chunk-1}\textbf{Table 1.} Quantitative and categorical traits used in this study.}
\centering
\fontsize{12}{14}\selectfont
\begin{tabular}[t]{ll}
\toprule
\textbf{Traits} & \textbf{Numerical range/Categories}\\
\midrule
\addlinespace[0.3em]
\multicolumn{2}{l}{\textbf{Quantitative}}\\
\hspace{1em}Plant height & 0.01-30m\\
\hspace{1em}Flower width & 0.50-205mm\\
\hspace{1em}Flower length & 0.20-195mm\\
\hspace{1em}Inflorescence width & 0.80-300mm\\
\hspace{1em}Style length & 0-138mm\\
\hspace{1em}Ovules per flower & 1-25000\\
\hspace{1em}Flowers per plant & 1-1.8 x $10^5$\\
\hspace{1em}Microliters of nectar & 0-160$\mu$l\\
\hspace{1em}Milligrams of nectar & 0-6.1mg\\
\hspace{1em}Pollen grains per flower & 13-2 x $10^8$\\
\hspace{1em}Autonomous selfing (fruit set) & 0-100$\%$\\
\addlinespace[0.3em]
\multicolumn{2}{l}{\textbf{Categorical}}\\
\hspace{1em}Lifepan & Short-lived and perennial\\
\hspace{1em}Life form & Herb, shrub and tree\\
\hspace{1em}Flower shape & \makecell[l]{Brush, campanulate, capitulum,\\ open, papilionaceous and tube}\\
\hspace{1em}Flower symmetry & Actinomorphic and zygomorphic\\
\hspace{1em}Nectar & Presence and absence\\
\hspace{1em}Compatibility system & \makecell[l]{Self-incompatible, partially self-compabtile \\ and self-compatible}\\
\hspace{1em}Breeding system & Hermaphrodite, monoecious and dioecious\\
\bottomrule
\end{tabular}
\end{table}

### 2023-02-14 Supplementary_Material/Sup_Mat_Lanuza_et_al_2022.Rmd
Edits on code to run without the last col of nectar conc that is not finally added

#Not included at the end
#Nectar mg
#nectar_mg <- readRDS("Data/RData/Nectar_mg_phylo.rds")
#Nectar concentration
nectar_con <- readRDS("Data/RData/Nectar_concentration_phylo.rds")

#Select two decimals function
specify_decimal <- function(x, k) trimws(format(round(x, k), nsmall=k))
#LABELS
label <- c("Autonomous selfing", "Flower number", "Inflorescence width" ,"Flower width", "Flower length", "Style length", "Ovule number", "Plant height", "Nectar per flower ($\\mu$l)", "Nectar concentration ($\\%$)","Pollen grains per flower")
#LAMBDA
lambda <- c(selfing$lambda , flower_n$lambda, inflow_width$lambda, flower_width$lambda, flower_length$lambda, style_n$lambda, ovule_n$lambda, plant_height$lambda, nectar_ul$lambda, nectar_con$lambda, pollen$lambda)
#P-VALUE
p_value <- c(selfing$P , flower_n$P, inflow_width$P,flower_width$P, flower_length$P, style_n$P, ovule_n$P, plant_height$P,nectar_ul$P,  nectar_con$P, pollen$P)

@@

#Select columns to work with
t <- trait_filtered_2[c("Species_geonet","Order_all","Family_all","Genus_all","Species_all","Breeding_system","Compatibility_system",
                        "Autonomous_selfing_level","Autonomous_selfing_level_fruit_set", "Flower_morphology", "Flower_symmetry", 
                        "Flowers_per_plant", "Floral_unit_width", "Corolla_diameter_mean", "Corolla_length_mean","Style_length", "Ovule_number", 
                        "life_form", "lifespan","Plant_height_mean_m","Nectar_presence_absence","Nectar_ul","Nectar_mg","Nectar_concentration",
                        "Pollen_per_flower")]

@@

#t$Nectar_ul adding it here just to see how many cells 
#Now we just select presence/absence of nectar
#select columns of interest
t <- t[c("Species_geonet","Order_all","Family_all","Genus_all","Species_all","Breeding_system","Compatibility_system",
         "Autonomous_selfing_level","Autonomous_selfing_level_fruit_set", "Flower_morphology", "Flower_symmetry", 
         "Flowers_per_plant", "Floral_unit_width", "Corolla_diameter_mean", "Corolla_length_mean","Style_length", "Ovule_number", 
         "life_form", "lifespan","Plant_height_mean_m","Nectar_presence_absence","Nectar_ul","Nectar_mg","Nectar_concentration",
         "Pollen_per_flower")]

@@

#Select just columns with traits
t <- t[c("Breeding_system","Compatibility_system",
                        "Autonomous_selfing_level","Autonomous_selfing_level_fruit_set", "Flower_morphology", "Flower_symmetry", 
                        "Flowers_per_plant", "Floral_unit_width", "Corolla_diameter_mean", "Corolla_length_mean","Style_length", "Ovule_number", 
                        "life_form", "lifespan","Plant_height_mean_m","Nectar_presence_absence","Nectar_ul","Nectar_mg","Nectar_concentration",
         "Pollen_per_flower")]

@@

colnames(t) <- c("Breeding system","Compatibility system","Autonomous selfing level", "Flower shape", "Flower symmetry", 
                 "Flower number per plant", "Inflorescence width", "Flower width", "Flower length","Style length", "Ovule number", 
                 "Life form", "Lifespan","Plant height","Nectar qualitative","Nectar per flower (microlitres)", "Nectar per flower (miligrames)", "Nectar concentration (%)", "Pollen grains per flower")

>>>>

#Not included at the end
#Nectar mg
#nectar_mg <- readRDS("Data/RData/Nectar_mg_phylo.rds")

#Select two decimals function
specify_decimal <- function(x, k) trimws(format(round(x, k), nsmall=k))
#LABELS
label <- c("Autonomous selfing", "Flower number", "Inflorescence width" ,"Flower width", "Flower length", "Style length", "Ovule number", "Plant height", "Nectar per flower ($\\mu$l)","Pollen grains per flower")
#LAMBDA
lambda <- c(selfing$lambda , flower_n$lambda, inflow_width$lambda, flower_width$lambda, flower_length$lambda, style_n$lambda, ovule_n$lambda, plant_height$lambda, nectar_ul$lambda, pollen$lambda)
#P-VALUE
p_value <- c(selfing$P , flower_n$P, inflow_width$P,flower_width$P, flower_length$P, style_n$P, ovule_n$P, plant_height$P,nectar_ul$P, pollen$P)

@@

#Select columns to work with
t <- trait_filtered_2[c("Species_geonet","Order_all","Family_all","Genus_all","Species_all","Breeding_system","Compatibility_system",
                        "Autonomous_selfing_level","Autonomous_selfing_level_fruit_set", "Flower_morphology", "Flower_symmetry", 
                        "Flowers_per_plant", "Floral_unit_width", "Corolla_diameter_mean", "Corolla_length_mean","Style_length", "Ovule_number", 
                        "life_form", "lifespan","Plant_height_mean_m","Nectar_presence_absence","Nectar_ul","Nectar_mg",
                        "Pollen_per_flower")]

@@

#t$Nectar_ul adding it here just to see how many cells 
#Now we just select presence/absence of nectar
#select columns of interest
t <- t[c("Species_geonet","Order_all","Family_all","Genus_all","Species_all","Breeding_system","Compatibility_system",
         "Autonomous_selfing_level","Autonomous_selfing_level_fruit_set", "Flower_morphology", "Flower_symmetry", 
         "Flowers_per_plant", "Floral_unit_width", "Corolla_diameter_mean", "Corolla_length_mean","Style_length", "Ovule_number", 
         "life_form", "lifespan","Plant_height_mean_m","Nectar_presence_absence","Nectar_ul","Nectar_mg",
         "Pollen_per_flower")]

@@

#Select just columns with traits
t <- t[c("Breeding_system","Compatibility_system",
                        "Autonomous_selfing_level","Autonomous_selfing_level_fruit_set", "Flower_morphology", "Flower_symmetry", 
                        "Flowers_per_plant", "Floral_unit_width", "Corolla_diameter_mean", "Corolla_length_mean","Style_length", "Ovule_number", 
                        "life_form", "lifespan","Plant_height_mean_m","Nectar_presence_absence","Nectar_ul","Nectar_mg",
         "Pollen_per_flower")]

@@

colnames(t) <- c("Breeding system","Compatibility system","Autonomous selfing level", "Flower shape", "Flower symmetry", 
                 "Flower number per plant", "Inflorescence width", "Flower width", "Flower length","Style length", "Ovule number", 
                 "Life form", "Lifespan","Plant height","Nectar qualitative","Nectar per flower (microlitres)", "Nectar per flower (miligrames)",  "Pollen grains per flower")
